---
title: "Lab1"
author: "Peter Omasta"
date: "1/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Finally get to do something in this lab! Yay! 

```{r, include=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)


```

Read in noaa_fisheries.csv data

```{r, include=FALSE}
landings <- read_csv(here::here("data","noaa_fisheries.csv")) %>% 
  janitor::clean_names() #cleaned up the names of the columns
```
Now going to make a tidy version of the data frame

```{r}
tidy_landing <- landings %>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_usd = parse_number(dollars_usd))
```

Just want to look at salmon information.

```{r}
salmon_landings <- tidy_landing %>% 
  mutate(afs_clean = str_remove(afs_name, pattern= "aggregate")) %>% 
  #Just want to keep the rows that contain the word salmon
  filter(str_detect(afs_clean, pattern = "salmon")) %>% 
  tidyr::separate(afs_clean, into = c("group", "species"))
```
Find some grouped summary data: 

find annual total US landings and dollar value(summing across all states) for each TYPE of salmon using `group_by()` + `summarize`.

```{r}
salmon_summary <- salmon_landings %>% 
  group_by(year, species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_usd)
  )
  
```

Make a graph! 

```{r}
salmon_graph <- ggplot(data = salmon_summary,
                       aes(x = year, y = tot_landings,
                           group = species)) +
  geom_line(aes(color = species))

salmon_graph
```

Now we will export our graph with ggsave 

```{r}
ggsave(plot = salmon_graph, 
       here::here("figures","salmon_graph.jpg"),
       height = 5,
       width = 8)
```

Make a nice table for the final 

```{r}

salmon_first_5 <- salmon_summary %>% 
  head(5)
```

Use kable to make pretty 

```{r}

```

